---
title: "The Data Science R Bootcamp Assignment"
author: "Sanja Cvetković"
date: "5/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## [Assignment Problem 1](http://tanjakec.github.io/Project_DSBC/AssignmentQs.html#Problem_1)

The first assignment is to load the delightful traffic accidents data from. Lets take a glimpse.

```{r include=FALSE}
#install.packages("vtree")
library(leaflet)
suppressPackageStartupMessages(library(lubridate))
library(vtree)
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyverse))

# read data
mydata <- read.csv("data/nez-opendata_02_19.csv", 
                   na.strings = c("","NA"),
                   stringsAsFactor = FALSE) 
dim(mydata)
# have a look at the first few rows
head(mydata)

# note that the variables are not named.
# name the variables
names(mydata) <- letters[1:7]
head(mydata)
str(mydata)
glimpse(mydata)

# separate date and time and save as two variables ('date' and 'time')
mydata <- separate(mydata, b, c("date", "time"), sep = ",")
glimpse(mydata)

summary(mydata)

# organise data: format as factors e, d and f columns 
cols <- c("e", "f", "g")
mydata[cols] <- lapply(mydata[cols], as.factor)
glimpse(mydata)
summary(mydata)

# format data and time columns
mydata$date <- format(as.POSIXct(mydata$date,format="%d.%m.%Y"),"%d/%m/%Y")
mydata$time <- format(as.POSIXct(mydata$time, format="%H:%M"),"%H:%M")
mydata$date <- as.Date(mydata$date, "%d/%m/%Y")
glimpse(mydata)



# change the names of the levels in the factor variable e
summary(mydata$e)
levels(mydata$e) <- c("material_demage", "deaths", "injured")
glimpse(mydata)

# change the names of the levels in the factor variable e
summary(mydata$f)
levels(mydata$f) <- c("one_vehicle", "two_vehicle_no_turn", "two_vehicle", "parked_vehicle", "pedestrian")
glimpse(mydata)

# rename the variables
mydata <- rename(mydata, id = a, long = c, lat = d, accident = e, type_acc = f, descrip = g)
```

```{r include=TRUE}
summary(mydata$type_acc)
summary(mydata$accident)
glimpse(mydata)
```

<h2>Which kind of the accident is the most frequent?</h2>

<h3>By type_acc feature</h3>
```{r include=TRUE}
mydata %>% 
  group_by(type_acc) %>% 
  ggplot(aes(x = type_acc, fill = type_acc)) +
        geom_bar() +
        ylab("Number of incidents") +
        xlab("Type of accident")
```
<h3>By accident feature</h3>
```{r include=TRUE}
mydata %>% 
  group_by(accident) %>% 
  ggplot(aes(x = accident, fill = accident)) +
        geom_bar() +
        ylab("Number of incidents") +
        xlab("Accident")
```

```{r include=FALSE}
vtree(mydata, "accident", palette = 3, sortfill = TRUE,  horiz = FALSE)
```

To conclude the max values by accident and type_acc are:
```{r include=TRUE}
mydata %>% group_by(accident) %>% summarize(Count = n()) %>% filter(Count == max(Count))
```

```{r include=TRUE}
mydata %>% group_by(type_acc) %>% summarize(Count = n()) %>% filter(Count == max(Count))
```

<h2>Which day in the week has the highest number of accidents?</h2>

First we add the weekday column

```{r include=TRUE}
mydata$weekday = weekdays(mydata$date)
glimpse(mydata)
```

Now we plot the incidents.

```{r include=TRUE}
mydata %>% 
  group_by(weekday) %>% 
  ggplot(aes(x = weekday, fill = weekday)) +
        geom_bar() +
        ylab("Number of incidents") +
        xlab("Weekday")
```

To conclude - the max incidents are:
```{r include=TRUE}
mydata %>% group_by(weekday) %>% summarize(Count = n()) %>% filter(Count == max(Count))
```
<h2>For each day in the week find the total number of each kind of accidents?</h2>

Let see the numbers
```{r include=TRUE}
my_summarize <- mydata %>% group_by(weekday, type_acc) %>% summarize(Count = n())
glimpse(my_summarize)
```
And now lets visualize

```{r include=TRUE}
ggplot(data = my_summarize, aes(x = weekday, y=Count, fill=type_acc)) +
      geom_bar(position="stack", stat="identity") +
      ylab("Number of incidents") +
      xlab("Weekday")
```

<h2>Identify time in the day (the hour) with the highest number of accidents</h2>

Lets add hour in day column:

```{r include=TRUE}
mydata$hour_in_day = substring(mydata$time, 0, 2)
glimpse(mydata)
mydata %>% group_by(hour_in_day) %>% summarise(Count = n())
```

Lets visualize it:

```{r include=TRUE}
mydata %>% 
  group_by(hour_in_day) %>% 
  ggplot(aes(x = hour_in_day, fill = hour_in_day)) +
        geom_bar() +
        ylab("Number of incidents") +
        xlab("Hour in day")
```

To conclude:

```{r include=TRUE}
mydata %>% group_by(hour_in_day) %>% summarize(Count = n()) %>% filter(Count == max(Count))
```

<h2>What is the most occurring type of the accident during the hour with the highest number of accidents?</h2>

Lets find the hour with the highest number of accidents.
```{r include=TRUE}
max_hour <- mydata %>% group_by(hour_in_day) %>% summarize(Count = n()) %>% filter(Count == max(Count))
glimpse(max_hour)
```

We're filtering the data - only the data of that single max hour.

```{r include=TRUE}
my_max_hour_data <- mydata %>% group_by(type_acc) %>% filter(hour_in_day == max_hour$hour_in_day)
glimpse(my_max_hour_data)
my_max_hour_data %>% group_by(type_acc) %>% summarize(Count = n()) %>% filter(Count == max(Count))
```

Now lets visualize.

```{r include=TRUE}
my_max_hour_data %>% 
  group_by(type_acc) %>% 
  ggplot(aes(x = type_acc, fill = type_acc)) +
        geom_bar() +
        ylab("Number of incidents") +
        xlab("Type of accident")
```
<h2>Creat a leafleta interactive map to pinpoint the accidents with the popup message of the description of the accidents.</h2>
This popup message should appear every time a user clicks on an accident marker.

```{r include=TRUE}

minlat <- min(mydata$lat)
maxlat <- max(mydata$lat)
minlng <- min(mydata$long)
maxlng <- max(mydata$long)

libs <- mydata %>% 
  leaflet() %>% 
  # add tiles to the leaflet object
  addTiles() %>%
  # setting the corners/centre of the map
  fitBounds(~minlng, ~minlat, ~maxlng, ~maxlat) %>% 
  # add the markers with the popups 
  addCircles(lng = ~long, lat = ~lat,
             radius = 150, weight = 10, color = "black",
             fillColor = "red", fillOpacity = 0.7,  
             popup = ~paste(descrip))
#print the map
libs
```

## [Assignment Problem 2](http://tanjakec.github.io/Project_DSBC/AssignmentQs.html#Problem_2)

